---
title: "California Fish and Wildlife Plots V2"
author: "Brooke Rose"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---
# Note: For the final dataset of cleaned plots, I decided to use species codes as the columns (instead of species names). I was having trouble reading the file with species full names. (I still saved a version with full species names)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r mac working directory, eval = FALSE}
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/brookerose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/brookerose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/brookerose/Google Drive/Franklin_grant/project/scripts/"
```

```{r franklin working directory}
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/scripts/"
wd$software <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/software/"
```

```{r loading libraries}
library("sf")
library("raster")
require(ggplot2)
require(flora) # for removing the species names from the authors' names
require(vroom) # read and write database
require(janitor) # for cleaning columns names (very useful for integrating databases later)
require(dplyr)
require(tidyr)
library(fuzzySim)
```

```{r extra data, echo = FALSE}
# Environmental data
env.stack <- stack(paste0(wd$output, 'predictors/env_stack/CFP_environmental_stack.grd')) # for coordinate reference system
cfp <- st_read(paste0(wd$data,"shapefiles/CFP/CFP_GIS.shp"))
cfp.trans <- st_transform(cfp, crs(env.stack))

us_states <- st_read(paste0(wd$data, "shapefiles/tl_2019_us_state/tl_2019_us_state.shp"))
california <- us_states %>%
  filter(NAME == "California")
```

# California Vegetation Survey Plots (from Rosie)
Received November 26, 2019. Includes a total of 25,449 plots

```{r Survey Plots}
survey_plots <- st_read(paste0(wd$data, 'BIOS/Cal_survey_points/Cal_survey_plots.shp'))
# Waypoint as character (for joining)
survey_plots$WyptID <- as.character(survey_plots$WyptID)
#str(survey_plots)
survey_plots <- survey_plots %>% janitor::clean_names()
```

# Plant List with Plot ID's (Waypoint)
List of plants at each plot, 461,718 "observations". Includes species name, code, species cover, and stratum (life form)

```{r plant list}
plant_list <- vroom::vroom(paste0(wd$data, 'BIOS/Cal_survey_points/SurveyPlants.csv'))
#str(plant_list)
plant_list <- plant_list %>% janitor::clean_names()
```

# Join survey plots and plant list
Creates a data frame with the waypoint ID, survey type, survey date, National Vegetation Classification Name, location information, species code, species name, and species cover

```{r joined dataframe}
survey_join <- left_join(
  survey_plots %>%
    dplyr::select(wypt_id, survey_type, survey_date, nvcs_name,
                  geometry),
  plant_list %>% dplyr::select(waypoint_id, code_species, species_name, species_cover),
  by = c("wypt_id" = "waypoint_id")
) %>%
  janitor::clean_names() %>%
  dplyr::mutate(x_coord = sf::st_coordinates(.)[, 1],
                y_coord = sf::st_coordinates(.)[, 2])
  

str(survey_join)

vroom::vroom_write(survey_join, paste0(wd$data, "BIOS/Cal_survey_points/plot_plant_join.gz"))

#st_write(survey_join, paste0(wd$data, "BIOS/Cal_survey_points/plot_plant_join.shp"), delete_layer = TRUE)
#write_csv(survey_join %>% dplyr::select(-geometry), paste0(wd$data, "BIOS/Cal_survey_points/plot_plant_join.csv"))
```


# Converting data to  presence absence matrix

```{r data cleaning}

survey_df <- vroom::vroom(paste0(wd$data, "BIOS/Cal_survey_points/plot_plant_join.gz"))

plots_info <- survey_df %>%
  dplyr::select(wypt_id, x_coord, y_coord, survey_type, survey_date) %>% unique

species_df <- survey_df %>%
  dplyr::select(wypt_id, code_species) %>% unique

# converting plot/species data frame to species presence and absences
pres_abs <- splist2presabs(
  species_df,
  sites.col = "wypt_id",
  sp.col = "code_species",
  keep.n = FALSE
)

# joining the plots with species presence and absence with other data
new_survey <- 
  left_join(plots_info, pres_abs, by = c("wypt_id")) 

filt <- new_survey %>% dplyr::select(-c(1:5)) %>% colnames %>% sort
#filt #vector with species symbol

# Database with accepted names
plant_codes_accepted <-
  plant_list %>% dplyr::filter(code_species %in% filt)

rm(plant_list)
```

#### Removing incorrect coordinates
```{r santiago cleaning code 2}

### Remove plot without coordinates
cat("Total number of plots: ", nrow(new_survey))
new_survey <- new_survey %>%
  dplyr::filter(complete.cases(new_survey %>% dplyr::select(x_coord, y_coord)))
cat("Number of plots with coordinates: ", nrow(new_survey))

### Remove plot with wrong coordinates
new_survey %>% ggplot(aes(x_coord, y_coord)) + 
  geom_hex(bins=200) + 
  coord_equal() + theme_minimal()

### Remove columns with species without any presence
filt <- new_survey %>% dplyr::select(c(1:5)) %>% names
pres_col <- colSums(new_survey %>% dplyr::select(-filt), na.rm = T)
pres_col <- pres_col[pres_col>0] %>% names

new_survey <- new_survey %>% dplyr::select(c(1:5), {pres_col}) 

# nrow(survey_data)
cat("Number of species with at least one presence point: ", length(pres_col))

### Fill with zero in those columns with species names
new_survey <-
  new_survey %>% 
  dplyr::mutate(dplyr::across(pres_col, ~ tidyr::replace_na(.x, 0)))

new_survey <- new_survey %>% arrange(wypt_id) #sort rows based on waypoint ID

### Remove  identical rows coordinates and summarize information at waypoint level 
new_survey %>% dplyr::count(wypt_id) %>% arrange(desc(n)) #almost all waypoint is repeated

df2 <- # unique data for each plot
  new_survey %>% dplyr::select(-c(pres_col)) %>% unique

df3 <- new_survey %>% # summarizing species records by plot 
  dplyr::select(c('wypt_id', c(pres_col))) %>%
  group_by(wypt_id) %>%
  summarise(across(c(pres_col), ~ sum(.x)))

new_survey2 <- dplyr::left_join(df2, df3, by = "wypt_id")
#dim(survey_data)
cat("Dimensions of clean data frame: ", dim(new_survey2))
new_survey <- new_survey2
rm(new_survey2)

nrow(unique(new_survey)) == nrow(new_survey) # all rows are unique. however there are plot with same coordinates 
# which not does not necessarily represent duplicate coordinates with different species composition

### Exploring plot without any species
rowfilt <- new_survey[pres_col] %>% colSums()
cat("All plots have at least one presence: ", range(rowfilt)) 
```

# Map of all survey plots

```{r maps}

study_sf <- st_as_sf(new_survey, coords = c("x_coord", "y_coord"), crs = 3310)  
study_sf <- st_transform(study_sf, st_crs(cfp.trans)) # data frame of presence/absence for Compadre species with shapefile capabilities

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = study_sf, aes(color = survey_type)) 

cat("Total number of plots: ", nrow(study_sf))
```

```{r data source}
cfw_sources <- study_sf %>%
  dplyr::mutate(
    survey = ifelse(
      survey_type == 'Transect' | survey_type == 'Releve' |
        survey_type == 'Multivisit Releve' |
        survey_type == 'releve' |
        survey_type == 'Multivisit Transect' |
        survey_type == 'Multi-visit releve' |
        survey_type == 'multi-visit transect' |
        survey_type == 'Relevee',
      'releve',
      'rapid'
    )
  )

cfw_cfp <- st_intersection(cfw_sources, st_make_valid(cfp.trans)) %>%
  dplyr::mutate(x_coords = sf::st_coordinates(.)[,1],
              y_coords = sf::st_coordinates(.)[,2])
cat("Total number of plots in the CFP: ", nrow(cfw_cfp))

vroom::vroom_write(cfw_cfp, paste0(wd$data, "BIOS/Cal_survey_points/cfw_final_cfp_sp_codes.gz"))
```

# Cal Fish & Wildlife Data by Survey Type
## Rapids

```{r rapids}
cfp_rapids <- cfw_cfp %>% filter(survey == 'rapid')
cat("Total number of rapid plots in the CFP: ", nrow(cfp_rapids))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = cfp_rapids, aes(color = survey_type)) 
```

## Relevees

```{r relevees}
cfp_releve <- cfw_cfp %>% filter(survey == 'releve')
cat("Total number of relevee plots in the CFP: ", nrow(cfp_releve))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = cfp_releve, aes(color = survey_type)) 
```