---
title: "Environmental Data Processing"
author: "Brooke Rose"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: journal
    highlight: espresso
bibliography: paperpile.bib
---

# Introduction
This Rmarkdown processes and compiles all of the environmental raster data that is used in the NSF Geography and Vulnerability Project. The final raster stack includes these variables for the extent of the California portion of the CFP:

1. Actual Evapotranspiration (aet) - BCM
2. Climatic Water Deficit (cwd) - BCM
3. Maximum Monthly Temperature (tmax) - BCM
4. Minimum Monthly Temperature (tmin) - BCM
5. Snowpack - BCM
6. Potential Evapotranspiration (pet) - BCM
7. Precipitation (precip) - BCM
8. Precipitation Dec, Jan, Feb (ppt_djf) - BCM
9. Precipitation June, July, Aug (ppt_jja) - BCM
10. Precipitation March, April, May (ppt_mam) - BCM
11. Precipitation Sep, Oct, Nov (ppt_son) - BCM
12. Recharge - BCM
13. Runoff - BCM
14. pH - gNATSGO
15. Available Water Holding Capacity (awc) - gNATSGO
16. Soil Depth (depth) - gNATSGO
17. Percent clay - gNATSGO
18. Landform Classes (ergo_landform) - Theobald 2015

I also create a version of the CFP shapefile that only includes the ecoregions within California, for simplicity (CFP_GIS_California_shp). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.fullwidth=TRUE)
```

```{r libraries, include = FALSE}
library(tidyverse)
theme_set(theme_bw()) # classic dark on white ggplot theme
library(sf)
library(maps)
library(ggrepel)
library(ggspatial)
library(lubridate)
library(rgdal)
library(leaflet)
library(raster)
library(rgdal)
library(rasterVis)
library(ggpubr)
library(extrafont)
library(ggcorrplot)
library(ggsn)
font_import()
loadfonts(device = "win")
```

```{r mac working directory, eval = FALSE}
# Setting directories
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/brookerose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/brookerose/Google Drive/Franklin_grant/project/output/"
```

```{r franklin working directory}
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/scripts/"
wd$software <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/software/"
wd$rmds <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/rmds/"
```

# Basin Charicterization Model 
Historical and projected climate and hydrology a 270 meter resolution, which is relevant for watershed-scale evaluation and planning
These data were downloaded on 11/11/19. Temperature units = Celcius and all other variables = mm. The website provides 30-year summary data for 3 historical/current time periods: 1921-1950; 1951-1980; and 1981-2010. The data area in NAD1983 California Teale Albers Equal Area Projection. 
Reference: Flint, L. E., A. L. Flint, J. H. Thorne, and R. Boynton. 2013. Fine-scale hydrologic modeling for regional landscape applications: the California Basin Characterization Model development and performance. Ecological Processes 2:1–21. 

Data are available at: (http://climate.calcommons.org/).

```{r raster crs}
# California Albers proj4string
ref_crs <- '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs'
```

## 1921-1950

```{r 1921-1950 climate, fig.fullwidth=TRUE}
aet21_50 <- raster(paste0(wd$data, 'climate/BCM_1921_1950/aet/aet1921_1950_ave_HST_1574209209.tif'))
cwd21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/cwd/cwd1921_1950_ave_HST_1574209832.tif'))
tmax21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/tmax/tmx1921_1950_ave_HST_1574210150.tif'))
tmin21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/tmin/tmn1921_1950_ave_HST_1574210195.tif'))
snowpack21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/snowpack/aprpck1921_1950_ave_HST_1574210095.tif'))
pet21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/pet/pet1921_1950_ave_HST_1574209957.tif'))
ppt21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/ppt/average/ppt1921_1950_ave_HST_1574209969.tif'))
ppt_djf21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/ppt/djf/ppt1921_1950djf_ave_HST_1595636546.tif'))
ppt_jja21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/ppt/jja/ppt1921_1950jja_ave_HST_1595636451.tif'))
ppt_mam21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/ppt/mam/ppt1921_1950mam_ave_HST_1595636507.tif'))
ppt_son21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/ppt/son/ppt1921_1950son_ave_HST_1595636421.tif'))
recharge21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/recharge/rch1921_1950_ave_HST_1574210048.tif'))
runoff21_50  <- raster(paste0(wd$data, 'climate/BCM_1921_1950/runoff/run1921_1950_ave_HST_1574210081.tif'))

# stacking 1951-1981 BCM data and writing stack to file
bcm_data21_50  <- stack(aet21_50 , cwd21_50 , tmax21_50 , tmin21_50 , snowpack21_50 , pet21_50 , ppt21_50 , ppt_djf21_50 , ppt_jja21_50 , ppt_mam21_50 , ppt_son21_50 , recharge21_50 , runoff21_50 )
names(bcm_data21_50) <- c('aet', 'cwd', 'tmax', 'tmin', 'snowpack', 'pet', 'ppt', 'ppt_djf', 'ppt_jja', 'ppt_mam', 'ppt_son', 'recharge', 'runoff')

crs(bcm_data21_50) <- ref_crs
writeRaster(bcm_data21_50 , paste0(wd$data,'climate/BCM_1921_1950/Stacks/bcm_1921_1950_stack.grd'), format = 'raster', overwrite = TRUE)
bcm_stack21_50<- stack(paste0(wd$data, 'climate/BCM_1921_1950/Stacks/bcm_1921_1950_stack.grd'))

plot(bcm_stack21_50)
stack21_50.df <- as.data.frame(bcm_stack21_50) %>%
  na.omit() 
corr <- round(cor(stack21_50.df), 1)
p.mat <- cor_pmat(stack21_50.df)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, title = "Correlation matrix of climate data 1921-1950")
```

## 1951-1980

```{r 1951-1980}
aet51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/aet/aet1951_1980_ave_HST_1574206944.tif'))
cwd51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/cwd/cwd1951_1980_ave_HST_1574207362.tif'))
tmax51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/tmax/tmx1951_1980_ave_HST_1574207734.tif'))
tmin51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/tmin/tmn1951_1980_ave_HST_1574208011.tif'))
snowpack51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/snowpack/aprpck1951_1980_ave_HST_1574207685.tif'))
pet51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/pet/pet1951_1980_ave_HST_1574207088.tif'))
ppt51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/ppt/average/ppt1951_1980_ave_HST_1574207384.tif'))
ppt_djf51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/ppt/djf/ppt1951_1980djf_ave_HST_1595635668.tif'))
ppt_jja51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/ppt/jja/ppt1951_1980jja_ave_HST_1595635792.tif'))
ppt_mam51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/ppt/mam/ppt1951_1980mam_ave_HST_1595635867.tif'))
ppt_son51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/ppt/son/ppt1951_1980son_ave_HST_1595636007.tif'))
recharge51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/recharge/rch1951_1980_ave_HST_1574207402.tif'))
runoff51_80 <- raster(paste0(wd$data, 'climate/BCM_1951_1981/runoff/run1951_1980_ave_HST_1574207655.tif'))

# stacking 1951-1981 BCM data and writing stack to file
bcm_data51_80 <- stack(aet51_80, cwd51_80, tmax51_80, tmin51_80, snowpack51_80, pet51_80, ppt51_80, ppt_djf51_80, ppt_jja51_80, ppt_mam51_80, ppt_son51_80, recharge51_80, runoff51_80)
names(bcm_data51_80) <- c('aet', 'cwd', 'tmax', 'tmin', 'snowpack', 'pet', 'ppt', 'ppt_djf', 'ppt_jja', 'ppt_mam', 'ppt_son', 'recharge', 'runoff')

crs(bcm_data51_80) <- ref_crs
writeRaster(bcm_data51_80, paste0(wd$data, 'climate/BCM_1951_1981/Stacks/bcm_1951_1981_stack.grd'), format = 'raster', overwrite = TRUE)
bcm_stack51_80 <- stack(paste0(wd$data, 'climate/BCM_1951_1981/Stacks/bcm_1951_1981_stack.grd'))

plot(bcm_stack51_80)

stack51_80.df <- as.data.frame(bcm_stack51_80) %>%
  na.omit() 
corr <- round(cor(stack51_80.df), 1)
p.mat <- cor_pmat(stack51_80.df)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, title = "Correlation matrix of climate data 1951-1980")
```

## 1981-2010

```{r 1981-2010 climate data}
aet81_10 <- raster(paste0(wd$data,'climate/BCM_1981_2010/aet/aet1981_2010_ave_HST_1573498620.tif'))
cwd81_10 <- raster(paste0(wd$data,'climate/BCM_1981_2010/cwd/cwd1981_2010_ave_HST_1573498984.tif'))
tmax81_10 <- raster(paste0(wd$data,'climate/BCM_1981_2010/tmax/tmx1981_2010_ave_HST_1573502276.tif'))
tmin81_10 <- raster(paste0(wd$data,'climate/BCM_1981_2010/tmin/tmn1981_2010_ave_HST_1573502157.tif'))
snowpack81_10 <- raster(paste0(wd$data,'climate/BCM_1981_2010/snowpack/aprpck1981_2010_ave_HST_1574288967.tif'))
pet81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/pet/pet1981_2010_ave_HST_1573502738.tif'))
ppt81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/ppt/average/ppt1981_2010_ave_HST_1573502574.tif'))
ppt_djf81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/ppt/djf/ppt1981_2010djf_ave_HST_1595634789.tif'))
ppt_jja81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/ppt/jja/ppt1981_2010jja_ave_HST_1595635039.tif'))
ppt_mam81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/ppt/mam/ppt1981_2010mam_ave_HST_1595634866.tif'))
ppt_son81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/ppt/son/ppt1981_2010son_ave_HST_1595635194.tif'))
recharge81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/recharge/rch1981_2010_ave_HST_1573502478.tif'))
runoff81_10 <- raster(paste0(wd$data, 'climate/BCM_1981_2010/runoff/run1981_2010_ave_HST_1573502361.tif'))


# stacking 1981-2010 BCM data and writing stack to file
bcm_data81_10 <- stack(aet81_10, cwd81_10, tmax81_10, tmin81_10, snowpack81_10, pet81_10, ppt81_10, ppt_djf81_10, ppt_jja81_10, ppt_mam81_10, ppt_son81_10, recharge81_10, runoff81_10)
names(bcm_data81_10) <- c('aet', 'cwd', 'tmax', 'tmin', 'snowpack', 'pet', 'ppt', 'ppt_djf', 'ppt_jja', 'ppt_mam', 'ppt_son', 'recharge', 'runoff')

crs(bcm_data81_10) <- ref_crs
writeRaster(bcm_data81_10, paste0(wd$data, 'climate/BCM_1981_2010/Stacks/bcm1981_2010_stack.grd'), format = "raster", overwrite = TRUE)
bcm_stack81_10 <- stack(paste0(wd$data, 'climate/BCM_1981_2010/Stacks/bcm1981_2010_stack.grd'))

plot(bcm_stack81_10)

stack81_10.df <- as.data.frame(bcm_stack81_10) %>%
  na.omit()
corr <- round(cor(stack81_10.df), 1)
p.mat <- cor_pmat(stack81_10.df)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, title = "Correlation matrix of climate data 1981-2010")
```

# California Floristic Province
These are the 8 ecoregions of the CFP that are fully within the state of California. 
```{r cfp and US, include = FALSE}
us.states <- st_read(paste0(wd$data, 'shapefiles/tl_2019_us_state/tl_2019_us_state.shp'))
us.trans <- st_transform(us.states, crs(bcm_stack81_10)) %>%
  filter(NAME == "Oregon" | NAME == "California" | NAME == "Nevada")

cfp <- st_read(paste0(wd$data,'shapefiles/CFP/CFP_GIS.shp'))
cfp.trans <- st_transform(cfp, crs(bcm_stack81_10))
cfp.bb <- st_as_sfc(st_bbox(cfp.trans))
cfp_california <- cfp.trans %>% 
  filter(JEP_REG != "Nevada CFP" & JEP_REG != "Baja CFP" & JEP_REG != "Oregon CFP")
st_write(cfp_california, paste0(wd$data,'shapefiles/CFP/CFP_GIS_California.shp'), delete_layer = TRUE)
```

```{r cfp map}
cfp.area <- ggplot() +
  geom_sf(data = us.trans, color = "black", fill = "papayawhip") +
  geom_sf(data = cfp.trans %>% filter(JEP_REG != 'Baja CFP' & JEP_REG != 'Oregon CFP' & JEP_REG != 'Nevada CFP'), aes(fill = JEP_REG)) +
  xlab("Longitude") + ylab("Latitude") +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.direction = "vertical", legend.position = "right", legend.key = element_blank(), 
        legend.background = element_rect(fill = "white"),
        text = element_text(size = 8, family = "serif", color = 'black'),
        axis.text.x = element_text(color = 'black', size = 8),
        axis.text.y = element_text(color = 'black', size = 8)) +
  scale_fill_brewer(type = "qual", palette = "Spectral") +
  scale_color_brewer(type = "qual", palette = "Spectral") +
  north(us.trans, location = 'topright', scale = .11) +
  ggsn::scalebar(cfp.trans, dist = 250, dist_unit = "km", transform = FALSE, st.size = 3, family = 'serif')
cfp.area

```

# Soil data
These were downloaded from the Gridded National Soil Survey Geographic Database gNATSGO database - includes pH, available water holding capacity, depth, and percent clay. These data were originally at a 10-meter resolution; however I resampled them to 270-meter resolution in ArcGIS using the toolbox and guide provided by gNATSGO. 

Reference: Soil Survey Staff. Gridded National Soil Survey Geographic (gNATSGO) Database for California. United States Department of Agriculture, Natural Resources Conservation Service. Available online at https://nrcs.app.box.com/v/soils. February 10, 2020 (FY2020 official release).

```{r soil data}
# Substrate data from the gNATSGO database
pH_270m  <- raster(paste0(wd$data, 'substrate/gNATSGO_CA_outputs/pH_270m_Lkup.tif'))
awc_270m <- raster(paste0(wd$data, 'substrate/gNATSGO_CA_outputs/AWC_270m_Lkup.tif'))
depth_270m <- raster(paste0(wd$data, 'substrate/gNATSGO_CA_outputs/Depth_270m_Lkup.tif'))
percent_clay <- raster(paste0(wd$data, 'substrate/gNATSGO_CA_outputs/soil_rasters/pct_clay.tif'))

soil_stack <- stack(pH_270m, awc_270m, depth_270m, percent_clay)

soil_proj <- projectRaster(soil_stack, bcm_stack81_10) #projecting to NAD1983 Teale Albers
plot(soil_proj)
writeRaster(soil_proj, paste0(wd$output, 'predictors/soil_stack270m.tif'), overwrite = TRUE)
```


# Theobald Landform data 
[@Theobald2015-um]

```{r landforms}
landform <- raster(paste0(wd$data, 'terrain/Ecologically_re/ergo_mosaic/ergo_mosaic_CFP.tif'))

landform_proj <- projectRaster(landform, bcm_stack81_10)
landform_round <- round(landform_proj)
landform_factor <- as.factor(landform_round)
landform_rat <- ratify(landform_factor, count = TRUE)
levels(landform_rat)[[1]] = cbind(levels(landform_rat)[[1]], 
                              landform = c("None", "Peak/ridge (warm)", "Peak/ridge", "Peak/ridge (cool)",
                                           "Mountain/divide", "Cliff/steep slope", "Upper slope (warm)", "Upper slope",
                                           "Upper slope (cool)", "Upper slope (flat)", "Lower slope (warm)", "Lower slope",
                                           "Lower slope (cool)", "Lower slope(flat)", "Valley", "Valley (narrow)"))
knitr::kable(levels(landform_rat))

plot(landform_rat)

writeRaster(landform_rat, paste0(wd$output, 'predictors/landforms.tif'), overwrite = TRUE)
```

# Geomorpho90m_v.1.0 (Amatulli et al. 2020)
These data are from Shared Google Drive made by the authors in folder geomorphometry_90m_wgs84
The naming scheme: variable abbreviation_resolution_DEM source layers_tiling system.format
Data are for the Geomorphological forms (geomorphons/geom) consist of 10 classes 
Based on pattern recognition (as opposed to differential geometry)
Features: 1-flat, 2-peak/summit, 3-ridge, 4-shoulder, 5-spur, 6-slope, 7-hollow, 8-footslope, 9-valley, 10-pit/depression. 
*Excluding these from the final raster stack because we are not using them in our models* 

```{r geomorpho}
geomorpho <- raster(paste0(wd$data, 'terrain/geomorpho90m_v.1.0/geomorphometry_90m_wgs84/mosaic/mosaic_raster.tif'))
geomorpho_proj <- projectRaster(geomorpho, bcm_stack81_10)
plot(geomorpho_proj)

writeRaster(geomorpho_proj, paste0(wd$output, 'predictors/geomorphons.tif'), overwrite = TRUE)
```

```{r env raster crop}
soil_proj <- stack(paste0(wd$output, 'predictors/soil_stack270m.tif'))
landform_rat <- raster(paste0(wd$output, 'predictors/landforms.tif'))

env_rasters <- list(soil_proj, landform_rat)

# Using lapply to crop and mask rasters
env_crops <- lapply(env_rasters, crop, y= cfp_california)
env_masks <- lapply(env_crops, raster::mask, cfp_california)
env_stack <- stack(env_masks)
env_stack[[5]] <- resample(env_stack[[5]], env_stack[[1]], method = 'ngb')
names(env_stack) <- c('pH', 'awc', 'depth', 'percent_clay', 'landform')

env_stack_proj <- projectRaster(env_stack, crs = '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs') 
#plot(env_stack_proj, nc = 3, nr = 2)
# This is the coordinate reference system for California Teal Alber's Equal Area Projection (differences between model 
# coefficients that use rasters with lat-long vs. equal area is small, but equal-area projections are recommended 
# by Budic, Didenko, and Dormann 2016 - Ecology and Evolution)
writeRaster(env_stack_proj, paste0(wd$output, 'predictors/env_stack/CFP_soil_terrain_stack.grd'), overwrite = TRUE)
writeRaster(env_stack_proj, paste0(wd$output, 'predictors/env_stack/CFP_soil_terrain_stack.tif'), overwrite = TRUE)
```

```{r climate raster crop}
climate_rasters <- list(bcm_stack21_50, bcm_stack51_80, bcm_stack81_10)
climate_crops <- lapply(climate_rasters, crop, y = cfp_california)
climate_masks <- lapply(climate_crops, raster::mask, cfp_california)

climate_proj <- lapply(climate_masks, projectRaster, crs = '+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs')

cfp21_50 <- stack(climate_proj[[1]], env_stack_proj)
cfp51_80 <- stack(climate_proj[[2]], env_stack_proj)
cfp81_10 <- stack(climate_proj[[3]], env_stack_proj)
```

# Environmental predictors to be used for SDM 
These rasters include the climate normals form 1981-2010. All of the data are at 270m resolution and cover the California portion of the CFP.

```{r fig.width=10, fig.height=20, fig.fullwidth=TRUE}
cfp81_10 <- stack(paste0(wd$output, 'predictors/env_stack/BCM1981_2010_CA_CFP_Stack.grd'))
raster::plot(cfp81_10, nc = 3, nr = 6)
```

```{r write rasters california cfp, eval = FALSE}
writeRaster(cfp21_50, paste0(wd$output, 'predictors/env_stack/BCM1921_1950_CA_CFP_Stack.grd'), overwrite = TRUE)
writeRaster(cfp51_80, paste0(wd$output, 'predictors/env_stack/BCM1951_1980_CA_CFP_Stack.grd'), overwrite = TRUE)
writeRaster(cfp81_10, paste0(wd$output, 'predictors/env_stack/BCM1981_2010_CA_CFP_Stack.grd'), overwrite = TRUE)
```

```{r write rasters}
# These rasters have the extent of the full CFP (somewhat inconsistent)
#writeRaster(cfp21_50, paste0(wd$output, 'predictors/env_stack/BCM1921_1950_CFP_Stack.grd'), overwrite = TRUE)
#writeRaster(cfp51_80, paste0(wd$output, 'predictors/env_stack/BCM1951_1980_CFP_Stack.grd'), overwrite = TRUE)
#writeRaster(cfp81_10, paste0(wd$output, 'predictors/env_stack/BCM1981_2010_CFP_Stack.grd'), overwrite = TRUE)
```

# References
