---
title: "Environmental Data + Plots"
author: "Brooke Rose"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: journal
    highlight: espresso
---

# Note: For the final dataset of cleaned plots, I decided to use species codes as the columns (instead of species names). I was having trouble reading the file with species full names. (I still saved a version with full species names)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Introduction
This Rmarkdown compiles the plot data from all of our sources (Thorne rapids, Thorne releves, and California Fish and Wildlife). I then use the environmental raster stack created in "Climate_processing.Rmd" to extract the values of each environmental predictor at the plot locations.

```{r mac working directory, eval = FALSE}
# working directory
wd <- list()

# commonly used paths in my working directory 
wd$data   <- "/Users/brookerose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/brookerose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/brookerose/Google Drive/Franklin_grant/project/scripts/"
wd$software <- "/Users/brookerose/Google Drive/Franklin_grant/project/software/"
wd$images <- "/Users/brookerose/Google Drive/Franklin_grant/project/images/"
```

```{r franklin working directory}
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/scripts/"
wd$software <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/software/"
```

```{r load libraries}
library(XML)
library(sf)
library(raster)
library(dismo)
library(rgdal)
library(maptools)
library(tidyverse)
library(rstudioapi)
library(ggcorrplot)
library(GGally)
library(mgcv)
library(gam)
library(ROCR)
library(rgbif)
library(formula.tools)
library(params)
library(rasterVis)
library(viridis)
library(cowplot)
library(rcartocolor)
library(caret)
library(biomod2)
library(MASS)
library(mgcv)
library(earth)
library(rpart)
library(mda)
library(Hmisc)
library(randomForest)
library(PresenceAbsence)
library(gridExtra)
library(grid)
library(kableExtra)
library(plyr)
library(dplyr)
library(vroom)
```

# Data Preparation
## 1. Importing environmental data 
For this project, I am using the coordinate reference system (CRS) from the Basin Characterization Model data. This can be changed later. The BCM data are the 30-year summaries for 1981-2010. These were processed/created in "Climate_processing.Rmd". The environmental data include:
1. Actual Evapotranspiration (aet) - BCM
2. Climatic Water Deficit (cwd) - BCM
3. Maximum Monthly Temperature (tmax) - BCM
4. Minimum Monthly Temperature (tmin) - BCM
5. Snowpack - BCM
6. Potential Evapotranspiration (pet) - BCM
7. Precipitation (precip) - BCM
8. Precipitation Dec, Jan, Feb (ppt_djf) - BCM
9. Precipitation June, July, Aug (ppt_jja) - BCM
10. Precipitation March, April, May (ppt_mam) - BCM
11. Precipitation Sep, Oct, Nov (ppt_son) - BCM
12. Recharge - BCM
13. Runoff - BCM
14. pH - gNATSGO
15. Available Water Holding Capacity (awc) - gNATSGO
16. Soil Depth (depth) - gNATSGO
17. Percent clay - gNATSGO
18. Landform Classes (ergo_landform) - Theobald 2015
19. Geomorpho - Amatulli et al. 2020

```{r environmental predictors}
env.stack <- stack(paste0(wd$output, 'predictors/env_stack/BCM1981_2010_CA_CFP_Stack.grd'))
names(env.stack)
plot(env.stack)
```

### Correlation matrix of environmental predictors

```{r}
env.stack.df <- as.data.frame(env.stack) %>%
  na.omit() %>%
  dplyr::select(-runoff, -recharge, -snowpack)
corr <- round(cor(env.stack.df), 1)
p.mat <- cor_pmat(env.stack.df)
ggcorrplot(corr, hc.order = TRUE,
    type = "lower", p.mat = p.mat)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)
```

## 2. California Floristic Province (CFP) shapefile for mapping and delineating eco-regions for modeling

```{r cfp}
cfp <- st_read(paste0(wd$data,'shapefiles/CFP/CFP_GIS_California.shp'))
cfp.trans <- st_transform(cfp, crs = proj4string(env.stack))
```

## 3. Plot data with presence absence records for study species 
Here, I compile the plot datasets provide by Jim Thorne and California Fish and Wildlife. Jim Thorne's original files are located in the data/Thorne_plots/ folder and include the 3 Rapid GIS CSV files. The Cal Fish and Wildlife data are in the data/BIOS/Cal_survey_points/ folder and are called "Cal_survey_plots.shp". These data were individually cleaned and processed in "Thorne_relevee_data.Rmd", "Thorne_rapids_data.Rmd", and the "Cal_Survey_Plots_Rosie.Rmd". Processing included converting each dataset to presence/abscence format, removing incorrect/duplicate coordinates and observations. Individual plot id's were retained in case we want information associated with the original plots. 

```{r compiling plot data}

Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 10)

cfw_plots <-
  vroom::vroom(paste0(wd$data, "BIOS/Cal_survey_points/cfw_final_cfp_sp_codes.gz")) %>%
  mutate(
    source = "Cal Fish and Wildlife",
    survey = ifelse(
      survey_type == 'Transect' |
        survey_type == 'Releve' |
        survey_type == 'Multivist Releve' |
        survey_type == 'Multivisit Transect' |
        survey_type == 'releve' |
        survey_type == "Multi-visit releve" |
        survey_type == "multi-visit transect" |
        survey_type == 'Relevee',
      'releve',
      'rapid'
    )
  )

thorne_rapids <- vroom::vroom(paste0(wd$data, "Thorne_plots/final_compiled_rapid_sp_codes.gz")) %>%
  mutate(
    source = "Jim Thorne",
    survey = 'rapid')

thorne_releves <-
  vroom::vroom(paste0(wd$data, "Thorne_plots/final_compiled_relevee_no_vtm.gz")) %>%
  mutate(
    source = "Jim Thorne",
    survey = 'releve'
    )


# amazing, time saving function for combining data frames with different columns
# adding coherent releve vs. rapid variable
all_plots <-
  dplyr::bind_rows(cfw_plots, thorne_rapids, thorne_releves)

all_plots <- all_plots[!duplicated(all_plots[c('x_coords', 'y_coords')]),] # removing duplicates

vroom::vroom_write(all_plots, paste0(wd$output, 'plots/compiled_plots.gz'))
```

```{r transforming data into spatial data object}
all_plots <- vroom::vroom(paste0(wd$output, 'plots/compiled_plots.gz'))

all_plots %>% ggplot(aes(x_coords, y_coords)) + 
  geom_hex(bins=200) + 
  coord_equal() + theme_minimal()

plots_sf <- all_plots %>%
  st_as_sf(coords = c('x_coords', 'y_coords'), crs = 3309, remove = FALSE)

plots_sf <- st_transform(plots_sf, crs = proj4string(env.stack))

ggplot() +
  geom_sf(data = cfp.trans, fill = "antiquewhite") +
  geom_sf(data = plots_sf, aes(color = survey), alpha = .5) +
  labs(title = "All plots in the CFP (50,306 plots)") +
  theme(title = element_text(color = 'black', size = 20, family = "serif"),
        text = element_text(size = 20, family = "serif", color = 'black')) +
  facet_wrap(~source)

ggsave(paste0(wd$data, 'plots/final_plots_map.jpg'))

```

## 4. Extracting environmental data values at plot locations and writing to file
```{r}
library(data.table)
sp.extract <- raster::extract(x = env.stack, y = plots_sf, df = TRUE, sp = TRUE, cellnumbers = TRUE)

sp.data <- st_as_sf(sp.extract, remove = FALSE) %>%
  st_transform(crs = proj4string(env.stack)) %>%
  st_set_geometry(NULL)

# removing observations that have NA for these environmental variables (they are generated by BCM and do not have 100% full coverage)

sp.env <- sp.data[ , c("aet", "cwd", "snowpack", "recharge", "runoff")]  

sp.env.full <- sp.data[complete.cases(sp.env), ] %>%
  tibble::rowid_to_column("ID") # adding unique id for all plots


vroom::vroom_write(sp.env.full, paste0(wd$data, 'plots/sp_pres_abs.gz'))
plots <- vroom::vroom(paste0(wd$data, 'plots/sp_pres_abs.gz'))
``` 


## 5. Test data frame for Santiago
This data frame includes presence/absence data for Abies magnifica, Queruc engelmanni, Pinus lambertiana, Calocedrus decurrens, and Pinus jeffreyi + extracted environmental variables.

```{r}
plots <- vroom::vroom(paste0(wd$data, 'plots/sp_pres_abs.gz'))

sp_subset <- plots %>%
  dplyr::select(
    ID,
    source,
    x_coords,
    y_coords,
    ABMA,
    QUEN,
    PILA,
    PIJE,
    CADE27,
    aet,
    cwd,
    tmax,
    tmin,
    snowpack,
    pet,
    ppt,
    ppt_djf,
    ppt_jja,
    ppt_mam,
    ppt_son,
    recharge,
    runoff,
    pH,
    awc,
    depth,
    percent_clay,
    landform
  )

vroom::vroom_write(sp_subset, paste0(wd$data, 'plots/test_sp_pres_abs.gz'))
```
