---
title: "Thorne Relevee Data"
author: "Brooke Rose"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: journal
    highlight: espresso
---

# Note: For the final dataset of cleaned plots, I decided to use species codes as the columns (instead of species names). I was having trouble reading the file with species full names. (I still saved a version with full species names)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r libraries}
library("sf")
library("raster")
require(ggplot2)
require(flora) # for removing the species names from the authors' names
require(vroom) # read and write database
require(janitor) # for cleaning columns names (very useful for integrating databases later)
require(dplyr)
require(tidyr)
```

```{r mac working directory}
# Setting directories
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/brookerose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/brookerose/Google Drive/Franklin_grant/project/output/"
```

```{r franklin working directory}
# working directory
wd <- list()

# commonly used paths in my working directory
wd$data   <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/data/"
wd$output <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/output/"
wd$scripts <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/scripts/"
wd$software <- "/Users/Brooke Rose/Google Drive/Franklin_grant/project/software/"
```

```{r extra data, echo = FALSE}
# Environmental data
env.stack <- stack(paste0(wd$output, 'predictors/env_stack/BCM1981_2010_CFP_Stack.grd')) # for coordinate reference system
cfp <- st_read(paste0(wd$data,"shapefiles/CFP/CFP_GIS.shp"))
cfp.trans <- st_transform(cfp, crs(env.stack))

us_states <- st_read(paste0(wd$data, "shapefiles/tl_2019_us_state/tl_2019_us_state.shp"))
california <- us_states %>%
  filter(NAME == "California")
```

# Exploring Jim Thorne's relevee dataset

```{r usda plant codes}
plant_codes <- vroom::vroom(paste0(wd$data, 'Thorne_plots/California USDA Plants codes.csv'))
str(plant_codes)
plant_codes <- plant_codes %>% janitor::clean_names()
```

## EPSG Code Descriptions
### 3309: NAD_1927_California_Teale_albers
### 32610: UTM zone 10N

## Relevee data

```{r relevee data}
#ac_releve <- read_csv(paste0(wd$data, 'Thorne_plots/Relevees_A_C.csv'))
#cp_releve <- read_csv(paste0(wd$data, 'Thorne_plots/Relevees_C_P.csv'))
#pz_releve <- read_csv(paste0(wd$data, 'Thorne_plots/Relevees_P_Z.csv'))

all_releves <-
  list.files(paste0(wd$data, 'Thorne_plots'),
             pattern = 'Relevees_',
             full.names = T)  %>%
  sapply(., vroom::vroom)
```

### Cleaning data
```{r santiago cleaning code 1}
all_releves <- dplyr::bind_rows(all_releves) 

filt <- all_releves %>% dplyr::select(-c(1:7)) %>% colnames %>% sort
#filt #vector with species symbol

# Database with accepted names
plant_codes_accepted <-
  plant_codes %>% dplyr::filter(symbol %in% filt)

# Database with accepted and synonyms names for those columns names with synonyms
plant_codes_synonym <-
  plant_codes %>% dplyr::filter(synonym_symbol %in% filt)
plant_codes_synonym <-
  plant_codes %>% dplyr::filter(symbol %in% plant_codes_synonym$symbol)
plant_codes_synonym <-
  plant_codes_synonym %>% filter(synonym_symbol %in% filt |
                                   is.na(synonym_symbol))

rm(plant_codes)
```

#### Removing incorrect coordinates
```{r santiago cleaning code 2}

### Clean columns names 
filt <- all_releves %>% dplyr::select(c(1:7)) %>% names
newnames <- all_releves[1, filt] %>% 
  janitor::clean_names() %>% names
colnames(all_releves)[1:7] <- newnames

### Remove plot without coordinates
cat("Total number of rapids plots: ", nrow(all_releves))
all_releves <- all_releves %>%
  dplyr::filter(complete.cases(all_releves %>% dplyr::select(lon, lat)))
cat("Number of rapids plots with coordinates: ", nrow(all_releves))

### Remove plot with wrong coordinates
all_releves %>% ggplot(aes(lon, lat)) + 
  geom_hex(bins=200) + 
  coord_equal() + theme_minimal()

### Remove columns with species without any presence
filt <- all_releves %>% dplyr::select(c(1:7)) %>% names
pres_col <- colSums(all_releves %>% dplyr::select(-filt), na.rm = T)
pres_col <- pres_col[pres_col>0] %>% names

all_releves <- all_releves %>% dplyr::select(c(1:7), {pres_col}) 
all_releves <- all_releves %>% dplyr::select(-old_plot_id)

# nrow(all_releves)
cat("Number of species with at least one presence point: ", length(pres_col))

### Fill with zero in those columns with species names
all_releves <-
  all_releves %>% 
  dplyr::mutate(dplyr::across(pres_col, ~ tidyr::replace_na(.x, 0)))

all_releves <- all_releves %>% arrange(new_id) #sort rows based on new_id

### Remove  identical rows coordinates and summarize information at new_id level 
all_releves %>% dplyr::count(new_id) %>% arrange(desc(n)) #almost all new_id is repeated

df2 <- # unique data for each plot
  all_releves %>% dplyr::select(-c('total_of_old_plot_id', pres_col)) %>% unique

df3 <- all_releves %>% # summarizing species records by plot 
  dplyr::select(c('new_id', c('total_of_old_plot_id', pres_col))) %>%
  group_by(new_id) %>%
  summarise(across(c('total_of_old_plot_id', pres_col), ~ sum(.x)))

all_releves2 <- dplyr::left_join(df2, df3, by = "new_id")
#dim(all_releves)
cat("Dimensions of clean data frame: ", dim(all_releves2))
all_releves <- all_releves2
rm(all_releves2)

# nrow(unique(all_releves)) == nrow(all_releves) # all rows are unique. however there are plot with same coordinates 
# which not does not necessarily represent duplicate coordinates with different species composition

### Exploring plot without any species
rowfilt <- all_releves[pres_col] %>% colSums()
cat("All plots have at least one presence: ", range(rowfilt)) # all plots have presences but ZZZD has 78,033
```

```{r data with species names}

all_releves_sp <- all_releves
filt <- names(all_releves_sp)[1:6]
pres_col <- names(all_releves_sp)[-c(1:6)]

#### Database with species names ###

for( i in 1:length(pres_col)){
  message('processing species name ', i)
  
  sp_symbol <- pres_col[i]
  if(sp_symbol %in% plant_codes_accepted$symbol) {
    sp_names <- plant_codes_accepted %>%
      dplyr::filter(symbol %in% sp_symbol) %>%
      pull(scientific_name_with_author)
    sp_names <- flora::remove.authors(sp_names)
    pres_col2 <- pres_col[i]
    names(pres_col2) <- sp_names
    all_releves_sp <- dplyr::rename(all_releves_sp, pres_col2)

  } else if (sp_symbol %in% plant_codes_synonym$synonym_symbol) {
    message('Updating species names')
    sp_symbol2 <- plant_codes_synonym %>%
      dplyr::filter(synonym_symbol %in% sp_symbol) %>% #note here is used synonym_symbol
      pull(symbol) #here is extracted the accepted names symbol
    sp_names <- plant_codes_synonym %>%
      dplyr::filter(symbol %in% sp_symbol2) %>%
      dplyr::filter(is.na(synonym_symbol)) %>%
      pull(scientific_name_with_author)

    sp_names <- flora::remove.authors(sp_names)
    pres_col2 <- pres_col[i]
    names(pres_col2) <- sp_names
    all_releves_sp <- dplyr::rename(all_releves_sp, pres_col2)
    
  } else {
    message('This species code has no species names in USDA database')
  }
}

vroom::vroom_write(all_releves_sp, 
                   paste0(wd$data, "Thorne_plots/Thorne_releves_cleaned_w_sp_names.gz"))

```

## Relevee Data
### What are the data sources for the relevee plots?
```{r data sources}
sarea <- c(unique(all_releves$sarea))
print(sarea)
```

### Number of plots by source
```{r notes on data sources}
source_data <- all_releves %>% group_by(sarea) %>% summarise(count=n())
source_data
```

### HARR 
#### These data are in lat lon

```{r harr plots}
harr <- all_releves %>% filter(sarea == "HARR")
harr_sf <- st_as_sf(harr, coords = c('lon', 'lat'), crs = st_crs(cfp))
harr_trans <- st_transform(harr_sf, crs(env.stack))
ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = harr_trans, aes(color = sarea))
```

### NAD83/UTM Zone 10N Plots
These plots use UTM Zone 10N EPSG code 32610. Includes data from USFSN, PRGG, POTT, CCBLM, and SFHREL. In the Veg_PlotInfo_Dec2009_for_Maki.xls file, CCBLM is designated as a rapid survey (not relevee).

```{r plots with epsg code 32610}
utm_plots <- all_releves %>% filter(sarea == 'USFSN' | sarea == "PRGG"|
                                     sarea == "POTT" | sarea == "CCBLM"| sarea == "SFHREL") %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 32610) %>%
  st_transform(crs(env.stack))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = utm_plots, aes(color = sarea))
```

### NAD27 California Albers Plots
These plots work with NAD27 California Albers and include data from VTM, BMC, and LPBO. 

```{r plots with epsg code 3309}
albers_plots <- all_releves %>% filter(sarea == "VTM" | sarea == 'BMC' | sarea == 'LPBO') %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 3309) %>%
  st_transform(crs(env.stack))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = albers_plots %>% filter(sarea != "VTM"), aes(color = sarea))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = albers_plots %>% filter(sarea == "VTM"), aes(color = sarea))
```

### WGS 84/UTM zone 11N (EPSG Code: 32611)
These plots seem to have EPSG Code 32611 (sources include SEKIAA, SEKIRA, and SEKIVM)

```{r other sources}
other_plots <- all_releves %>% filter(sarea == 'SEKIAA'| sarea == 'SEKIRA' | sarea == 'SEKIVM') %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 32611) %>%
  st_transform(crs(env.stack))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = other_plots, aes(color = sarea))

other_df <- all_releves %>% filter(sarea == 'SEKIAA'| sarea == 'SEKIRA' | sarea == 'SEKIVM')
```

### SAMO
Seems inconsistent with the other unknown data. They look very similar to the UTM data. Also, in the Veg_PlotInfo_Dec2009_for Maki.xls file, these are labeled as rapids (not relevees). 2,606 plots.

```{r samo plots}
samo <- all_releves %>% filter(sarea == 'SAMO') %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 32611) %>%
  st_transform(crs(env.stack)) %>%
  filter(new_id != 'SAMOrap0027')  # from looking at the map in ArcGIS, this plot is pretty far east of the Santa Monica Mountains (it's also the only plot there and has no "presences" so I decided to exclude it to be safe)

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = samo, aes(color = sarea))

#st_write(samo_plots, paste0(wd$data, "Thorne_plots/samo_plots.shp"), delete_layer = TRUE)
#st_write(samo_plots, paste0(wd$data, "Thorne_plots/samo_plots.csv"), delete_layer = TRUE)
```

#### SAMO plots in the CFP
Here, I select the SAMO data that are within the CFP (2,520 plots). This is the vast majority of the data. Minus the plot that is to the east, the other data line up well with the Santa Monica Mountains. I excluded one odd plot that was pretty far east of the Santa Monica Mountains (NewID = SAMOrap0027) - found using ArcGIS. 

```{r samo plots in cfp}
samo_cfp <- samo %>% 
  st_intersection(st_make_valid(cfp.trans)) # from looking at the map in ArcGIS, this plot is pretty far east of the Santa Monica Mountains (it's also the only plot there and has no "presences" so I decided to exclude it to be safe)

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = samo_cfp, aes(color = sarea))
```

### SEKINRI
Seems inconsistent with other known data CRS. The longitude values are in the 300's and the latitude values are in the 4,000's. I am thinking there are some missing digits in these data. 626 plots. FIXED: I ended up multiplying longitude and latitude by 1000. The data now match the other Sequoia/Kings Canyon data. 

```{r sekinri plots}
# original data
sekinri_og <- all_releves %>% filter(sarea == 'SEKINRI') %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 32611) %>%
  st_transform(crs(env.stack)) %>%
  mutate(status = 'original_data')


sekinri_new <- all_releves %>% filter(sarea == 'SEKINRI') %>%
   mutate(new_lat = lat *1000,
         new_lon = lon *1000) %>%
  st_as_sf(coords = c('new_lon', 'new_lat'), crs = 32611) %>%
  st_transform(crs(env.stack)) 

names(sekinri_new)[names(sekinri_new) == "lat"] <- 'old_lat'
names(sekinri_new)[names(sekinri_new) == "lon"] <- 'old_lon'
  
sekinri_new <- sekinri_new %>%
  dplyr::select(-old_lat, -old_lon)

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = sekinri_og, aes(color = 'blue')) +
  geom_sf(data = sekinri_new, aes(color = 'orange'))



```


### Integrating the plots
Here I integrate the plots and filter by location and species code. The final data include our study species with plots that are located within the California Floristic Province.

```{r}
int_plots <- dplyr::bind_rows(albers_plots, harr_trans, utm_plots, other_plots, samo, sekinri_new)

relevee_sp <- int_plots %>%
  st_as_sf() %>%
  dplyr::mutate(x_coords = sf::st_coordinates(.)[,1],
              y_coords = sf::st_coordinates(.)[,2])

relevee_sp <- st_intersection(relevee_sp, st_make_valid(cfp.trans))
cat("Number of relevee plots in CFP (including VTM): ", nrow(relevee_sp))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = relevee_sp, aes(color = sarea))

vroom::vroom_write(relevee_sp, paste0(wd$data, "Thorne_plots/compiled_relevee_with_vtm.gz"))
```

### Removing VTM plots
These are from the 1930's, maybe we will use them for another study. 
```{r}
relevee_novtm <- relevee_sp %>% filter(sarea != 'VTM') %>%
dplyr::mutate(x_coords = sf::st_coordinates(.)[,1],
              y_coords = sf::st_coordinates(.)[,2])
cat("Number of relevee plots in CFP (minus VTM): ", nrow(relevee_novtm))

ggplot() +
  geom_sf(data = cfp.trans) +
  geom_sf(data = relevee_novtm, aes(color = sarea))

vroom::vroom_write(relevee_novtm, paste0(wd$data, "Thorne_plots/final_compiled_relevee_no_vtm.gz"))

```
